---
title: "Tier Creation for Publication"
output:
  pdf_document: default
  html_notebook: default
---
# The following code will create the tier (by following if then statements guided by the outlines given by the CEEW report) in 2015 and then 2018. The code will then find the change in tier between these two years for each observation. Then, if this change is greater than 1, a binary variable for "moving up a tier" will be assigned. By household id, these values will then be added as a column to the appended data set and a csv file will be created for further analysis in STATA. 

#recode household IDs for 2015 and 2018 
```{r - create new variables for 2015 and 2018 ids }
CEEW_ACCESS20152018_Merged$HHID15 <- CEEW_ACCESS20152018_Merged$m1_q3_hhid
CEEW_ACCESS20152018_Merged$HHID18 <- CEEW_ACCESS20152018_Merged$tm1_q3_hhid

```

# Created Tiers in the Merged Data for 2015
```{r - generate tiers for the Merged DATA 2015}
#tier 3
CEEW_ACCESS20152018_Merged$tier3part1 <- recode(CEEW_ACCESS20152018_Merged$m4_q103_lpg, "0=0; 1=3") 
CEEW_ACCESS20152018_Merged$tier3part2 <- recode(CEEW_ACCESS20152018_Merged$m4_q109_firewood, "0=3; 1=0") 
CEEW_ACCESS20152018_Merged$tier3part3 <- recode(CEEW_ACCESS20152018_Merged$m4_q113_dungcake, "0=3; 1=0")
CEEW_ACCESS20152018_Merged$tier3part4 <- recode(CEEW_ACCESS20152018_Merged$m4_q114_agro, "0=3; 1=0")
CEEW_ACCESS20152018_Merged$tier3part5 <- recode(CEEW_ACCESS20152018_Merged$m4_q115_other_fuel, "0=3; 1=0")
CEEW_ACCESS20152018_Merged$tier3total <- CEEW_ACCESS20152018_Merged$tier3part1+CEEW_ACCESS20152018_Merged$tier3part1+CEEW_ACCESS20152018_Merged$tier3part2+CEEW_ACCESS20152018_Merged$tier3part3+CEEW_ACCESS20152018_Merged$tier3part4+CEEW_ACCESS20152018_Merged$tier3part5

CEEW_ACCESS20152018_Merged$tier3 <- ifelse(CEEW_ACCESS20152018_Merged$tier3total==15, 4,0)
#overalltier
CEEW_ACCESS20152018_Merged$tier <- ifelse(CEEW_ACCESS20152018_Merged$m4_q103_lpg==0,0,0)
CEEW_ACCESS20152018_Merged$tier <- ifelse((CEEW_ACCESS20152018_Merged$m4_q103_lpg==1 && CEEW_ACCESS20152018_Merged$m5_q119_main_cookfuel_avail>=2 && CEEW_ACCESS20152018_Merged$m5_q121_cook_difficult==0 && CEEW_ACCESS20152018_Merged$m5_q121_cook_expensive==0 && CEEW_ACCESS20152018_Merged$m5_q121_cook_quality==1 && CEEW_ACCESS20152018_Merged$m5_q121_cook_muchtime==0), 2, CEEW_ACCESS20152018_Merged$tier) 

CEEW_ACCESS20152018_Merged$tier <-                                      ifelse((CEEW_ACCESS20152018_Merged$m4_q103_lpg==1 && CEEW_ACCESS20152018_Merged$m5_q119_main_cookfuel_avail<=1 | CEEW_ACCESS20152018_Merged$m5_q121_cook_difficult==1 | CEEW_ACCESS20152018_Merged$m5_q121_cook_expensive==1 | CEEW_ACCESS20152018_Merged$m5_q121_cook_quality==0| CEEW_ACCESS20152018_Merged$m5_q121_cook_muchtime==0),1, CEEW_ACCESS20152018_Merged$tier)
                                                                            CEEW_ACCESS20152018_Merged$tier <-                                                                  ifelse(CEEW_ACCESS20152018_Merged$tier3==4,3,CEEW_ACCESS20152018_Merged$tier)
```

#Create tiers for Merged Data 2018
```{r - generate tiers for the Merged DATA 2018}
#tier 3
CEEW_ACCESS20152018_Merged$tier3part118 <- recode(CEEW_ACCESS20152018_Merged$tm4_q103_lpg, "0=0; 1=3") 
CEEW_ACCESS20152018_Merged$tier3part218 <- recode(CEEW_ACCESS20152018_Merged$tm4_q109_firewood, "0=3; 1=0") 
CEEW_ACCESS20152018_Merged$tier3part318 <- recode(CEEW_ACCESS20152018_Merged$tm4_q113_dungcake, "0=3; 1=0")
CEEW_ACCESS20152018_Merged$tier3part418 <- recode(CEEW_ACCESS20152018_Merged$tm4_q114_agro, "0=3; 1=0")
CEEW_ACCESS20152018_Merged$tier3part518 <- recode(CEEW_ACCESS20152018_Merged$tm4_q115_other_fuel, "0=3; 1=0")
CEEW_ACCESS20152018_Merged$tier3total18 <- CEEW_ACCESS20152018_Merged$tier3part118+CEEW_ACCESS20152018_Merged$tier3part118+CEEW_ACCESS20152018_Merged$tier3part218+CEEW_ACCESS20152018_Merged$tier3part318+CEEW_ACCESS20152018_Merged$tier3part418+CEEW_ACCESS20152018_Merged$tier3part518

CEEW_ACCESS20152018_Merged$tier318 <- ifelse(CEEW_ACCESS20152018_Merged$tier3total18==15, 4,0)
#overall tiers
CEEW_ACCESS20152018_Merged$tier18 <- ifelse(CEEW_ACCESS20152018_Merged$tm4_q103_lpg==0,0,0)
CEEW_ACCESS20152018_Merged$tier18 <- ifelse((CEEW_ACCESS20152018_Merged$tm4_q103_lpg==1 && CEEW_ACCESS20152018_Merged$tm5_q119_main_cookfuel_avail>=2 && CEEW_ACCESS20152018_Merged$tm5_q121_cook_difficult==0 && CEEW_ACCESS20152018_Merged$tm5_q121_cook_expensive==0 && CEEW_ACCESS20152018_Merged$tm5_q121_cook_quality==1 && CEEW_ACCESS20152018_Merged$tm5_q121_cook_muchtime==0), 2, CEEW_ACCESS20152018_Merged$tier18) 

CEEW_ACCESS20152018_Merged$tier18 <-                                      ifelse((CEEW_ACCESS20152018_Merged$tm4_q103_lpg==1 && CEEW_ACCESS20152018_Merged$tm5_q119_main_cookfuel_avail<=1 | CEEW_ACCESS20152018_Merged$tm5_q121_cook_difficult==1 | CEEW_ACCESS20152018_Merged$tm5_q121_cook_expensive==1 | CEEW_ACCESS20152018_Merged$tm5_q121_cook_quality==0| CEEW_ACCESS20152018_Merged$tm5_q121_cook_muchtime==0),1, CEEW_ACCESS20152018_Merged$tier18)
                                                                            CEEW_ACCESS20152018_Merged$tier18 <-                                                                  ifelse(CEEW_ACCESS20152018_Merged$tier318==4,3,CEEW_ACCESS20152018_Merged$tier18)
```

#Create delta tiers for Merged Data
```{r - generate a variable for delta tiers}
CEEW_ACCESS20152018_Merged$deltatier<- CEEW_ACCESS20152018_Merged$tier18-CEEW_ACCESS20152018_Merged$tier
CEEW_ACCESS20152018_Merged$moveuptier <- ifelse(CEEW_ACCESS20152018_Merged$deltatier>=1,1,0)
stat.desc(CEEW_ACCESS20152018_Merged$moveuptier)
Mergedonlyboth <- CEEW_ACCESS20152018_Merged[-c(which(is.na(CEEW_ACCESS20152018_Merged$deltatier))),]
tempmerg <- Mergedonlyboth[ ,c("HHID18","tier","tier18","deltatier")]
tempmerggathered <- tempmerg %>% gather("Tiertype","deltatier",tier,tier18) %>% mutate(year=ifelse(Tiertype=="tier", 0,1))
tempmerggathered <- tempmerggathered[,-2]
colnames(tempmerggathered)<- c("m1_q3_hhid","deltatier","AFT")
tempmerggathered<-distinct(tempmerggathered) 

TierAppendDF<-full_join(CEEW_ACCESS20152018_Appended,tempmerggathered, by=c("m1_q3_hhid","AFT"), all.x=TRUE,all.y=TRUE)


 TierAppendDF$moveuptier <- ifelse(TierAppendDF$deltatier>=1,1,0)
 
 write.csv(TierAppendDF,"C:\\Users\\agillwiehl\\Desktop\\TierAppendDF.csv", row.names=FALSE)
```
 
 

